---
title: "Project5"
author: "Jackson Dial"
date: "3/28/2022"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidyselect)
library(readxl)
```

```{r}
# dat_t <- read_csv("Project1/TransposedData.csv")
# 
# 
# 
# #, `strata_sex_label.Deaths from All Causes` == "Both"
# 
# testtv <- dat_t %>%
#   select(vars_select(names(dat_t), starts_with(c('value.', 'geo_label_city', 'date_label', 'strata_race_label', 'strata_sex_label', 'strata_race_sex'), ignore.case = TRUE))) %>% 
#   filter(`strata_race_label.Deaths from All Causes` == "All") %>% 
#   select(value.Suicide, `value.Motor Vehicle Deaths`, `strata_sex_label.Suicide`, strata_race_sex_label.Suicide)

#  separate(strata_race_sex_label.Suicide, into = c("race2", "sex2"), sep = ' ') %>% 


megan_dat <- read_xlsx("Project1/megan_dat.xlsx") 
megan_dat[megan_dat==""] <- NA


final_dat <- megan_dat %>% 
  filter(strata_race_label=="All") %>%
  filter(strata_sex_label!="Both") %>%
  filter(is.na(strata_race_sex_label)) %>%
  filter(metric_item_label=="Motor Vehicle Deaths" | metric_item_label=="Suicide") %>%
  select(metric_item_label,geo_label_citystate,value,date_label,strata_sex_label) %>%
  pivot_wider(names_from = metric_item_label,values_from = value) %>% 
  rename(mvd = `Motor Vehicle Deaths`,
         sex = strata_sex_label)

table(final_dat$sex)
```


```{r}
fit1 <- aov(lm(Suicide ~ (as.factor(sex) + mvd), data = final_dat))
summary(fit1)
```

After adjustment for Motor Vehicle Deaths, there is a significant difference in suicide rates between sex.

```{r}
1 - pf(1864.6, 1, 577)
```

```{r}
ggplot(final_dat, aes(x = mvd, y = Suicide, color = as.factor(sex)))+
  geom_point()+
  theme(panel.grid.minor = element_blank())+
  labs(x = "Motor Vehicle Deaths",
       y = "Suicide Rates",
       title = "Association between Suicide and MVD by Sex",
       color = "Sex")

ggplot(final_dat, aes(x = as.factor(sex), y = Suicide, fill = as.factor(sex)))+
  geom_boxplot()+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(x = "Sex",
       y = "Suicide Rates",
       title = "Boxplots of Suicide by Sex",
       fill = "Sex")

ggplot(final_dat, aes(x = as.factor(sex), y = mvd, fill = as.factor(sex)))+
  geom_boxplot()+
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())+
  labs(x = "Sex",
       y = "Motor Vehicle Deaths",
       title = "Boxplots of MVD by Sex",
       fill = "Sex")

```

### Shapiro test for assumptions of correlation test

```{r}
shapiro.test(final_dat$Suicide)
shapiro.test(final_dat$mvd)
```

```{r}
cor.test(final_dat$Suicide, final_dat$mvd, method = "pearson")
```



```{r}
fit2 <- lm(Suicide ~ as.factor(sex) + mvd, data = final_dat)
summary(fit2)

plot(fit2)
```

### Just male

```{r}
male_dat <- final_dat %>% filter(sex == "Male")
fit_male <- lm(Suicide ~ mvd, data = male_dat)
summary(fit_male)
```

```{r}
plot(fit_male)
shapiro.test(male_dat$Suicide)
shapiro.test(male_dat$mvd)

```

Since the shapiro wilks test concludes non-normality, I will perform log transformation:

```{r}
logged_dat <- male_dat %>% transmute(
  suicide = log(Suicide),
  mvd = log(mvd))

fit_male_log <- lm(suicide ~ mvd, data = logged_dat)
plot(fit_male_log)
shapiro.test(logged_dat$suicide)
shapiro.test(logged_dat$mvd)
```


Because we still have non-normality, I am going to run a boxCox test to see if log is the best transformation.

```{r}
library(car)
boxCox(fit_male)
```

Since 0 is within the bounds, the log transformation was the optimal choice. Guess we just won't fully meet our assumptions.

